---
title: "R Notebook"
output: html_notebook
---

Including Libraries

```{r}
library(readr)
library(dplyr)
```

Read CSV files of pre and post

```{r}
pre_data <- read_csv('../TLM_Exp_1_PRE.csv', col_types = cols(
  uid = col_character(),
  group = col_factor()
) )

post_data <- read_csv('../TLM_Exp_1_POST.csv', col_types = cols(
  uid = col_character(),
  group = col_factor()
))
```

```{r}
print(levels(pre_data$group))
```

```{r}
print(levels(post_data$group))
```

```{r}
pre_data %>% select(uid, group) %>% summary()
```

```{r}
post_data %>% select(uid, group) %>% summary()
```

Identify rows with missing values -- PRE

```{r}
rows_with_missing <- which(rowSums(is.na(pre_data)) > 0)
rows_with_missing
```

Print rows with missing values PRE

```{r}
print(pre_data[rows_with_missing, ])
# ONLY 5 STUDENTS HAVE MISSING DATA
```

Replace NA with 0

```{r}
pre_df_nona <- pre_data %>% mutate_at(vars(-c(uid, group)), ~replace(., is.na(.), 0))

```

Identify rows with missing values -- PRE NEW

```{r}
rows_with_missing <- which(rowSums(is.na(pre_df_nona)) > 0)
rows_with_missing
```

## Repeat the same NA identification and replacement on Post data

Identify rows with missing values -- post

```{r}
rows_with_missing <- which(rowSums(is.na(post_data)) > 0)
rows_with_missing
```

Print rows with missing values post

```{r}
print(post_data[rows_with_missing, ])
# ONLY 5 STUDENTS HAVE MISSING DATA
```

Replace NA with 0

```{r}
post_df_nona <- post_data %>% mutate_at(vars(-c(uid, group)), ~replace(., is.na(.), 0))

```

Identify rows with missing values -- post NEW

```{r}
rows_with_missing <- which(rowSums(is.na(post_df_nona)) > 0)
rows_with_missing
```

```{r}

## PRE
# calculate the total duration by summing duration in all subcategories
pre_complete <- pre_df_nona %>% rowwise() %>%
    mutate(total_time = sum(c(duration_aa, duration_as, duration_dd, duration_in, duration_ii), na.rm = TRUE))

# calcualte total score by summing the score of each subcategory
pre_complete <- pre_complete %>% rowwise() %>%
mutate(score = sum(c(total_aa, total_as, total_dd, total_in, total_ii), na.rm = TRUE))

pre_complete$score
# create a new variable containing only these columns
pre_totals <- pre_complete %>% select(uid, group, total_aa, total_as, total_dd, total_in, total_ii, total_time, score)
```

```{r}
# show it
pre_totals
```

```{r}

# show the mean of both groups at PRE for final total score
pre_totals %>% group_by(group) %>% summarise(mean_total=mean(score), .groups = 'drop')
# show the mean of both groups at PRE for each subcateory, time, and final total score
pre_totals %>% select( group, total_aa, total_as, total_dd, total_in, total_ii, total_time, score) %>% 
              group_by(group) %>% summarise(across(everything(), mean))
```

```{r}
## POST
# calculate the total duration by summing duration in all subcategories
post_complete <- post_df_nona %>% rowwise() %>%
  mutate(total_time = sum(c(duration_aa, duration_as, duration_dd, duration_in, duration_ii), na.rm = TRUE))

# calcualte total score by summing the score of each subcategory
post_complete <- post_complete %>% rowwise() %>%
  mutate(score = sum(c(total_aa, total_as, total_dd, total_in, total_ii), na.rm = TRUE))

# create a new variable con/taining only these columns
post_totals <- post_complete %>% select(uid, group, total_aa, total_as, total_dd, total_in, total_ii, total_time, score)
```

```{r}
# show it
post_totals
```

```{r}
# show the mean of both groups at post for final total score
post_totals %>% group_by(group) %>% summarise(mean_total=mean(score), .groups = 'drop')
# show the mean of both groups at post for each subcateory, time, and final total score
post_totals %>% select( group, total_aa, total_as, total_dd, total_in, total_ii, total_time, score) %>% 
  group_by(group) %>% summarise(across(everything(), mean),
                                .groups = 'drop')  %>% as.data.frame()

```

### prepare data to calculate the normalized learning gain

```{r}
names(pre_totals)
```

```{r}
names(pre_totals) <- c('uid', 'group', 'pre_total_aa', 'pre_total_as', 'pre_total_dd', 'pre_total_in', 'pre_total_ii', 'pre_total_time', 'pre_score' )
names(post_totals) <- c('uid', 'group', 'post_total_aa', 'post_total_as', 'post_total_dd', 'post_total_in', 'post_total_ii', 'post_total_time', 'post_score') 
```

```{r}
# Merge the pre-test and post-test datasets based on the user id
pre_post_totals <- merge(pre_totals, post_totals, by = c("uid", 'group'), all = FALSE)
nrow(pre_post_totals)
```

```{r}
names(pre_post_totals)

```

```{r}
pre_post_totals %>% group_by(group) %>% count(group)
```

calculate normalized gain for score Normalized Gain = (post_score - pre_score) / (maximum possible score - pre_score)

```{r}
# create the dataframe
pre_post_ng_score <- pre_post_totals %>% select(uid, group, pre_score, post_score)
pre_post_ng_score
```

calculate the difference

```{r}
pre_post_ng_score <- pre_post_ng_score %>% mutate(diff=post_score-pre_score)
```

calculate normalized gain

```{r}
pre_post_ng_score <- pre_post_ng_score %>% mutate(ng=diff/(40-pre_score))
pre_post_ng_score$ng <- round(pre_post_ng_score$ng, 3)
pre_post_ng_score
```

#### show the comparison of normalized gain between control and experiment

```{r}
pre_post_ng_score %>% group_by(group) %>%
  summarize(mean_ng = mean(ng),
            sd_ng = sd(ng))
```

#### show the comparison of total time spent on the test between control and experiment

```{r}
pre_post_totals %>% group_by(group) %>%
  summarize(mean_pre_total_time = mean(pre_total_time),
            sd_pre_total_time = sd(pre_total_time),
            mean_post_total_time = mean(post_total_time),
            sd_post_total_time = sd(post_total_time))
```

#### calculate time spent per category

```{r}
pre_times <- pre_df_nona %>% select(uid, group, duration_aa, duration_as, duration_dd, duration_in, duration_ii)
names(pre_times) = c('uid', 'group', 'pre_duration_aa', 'pre_duration_as', 'pre_duration_dd', 'pre_duration_in', 'pre_duration_ii')
post_times <- post_df_nona %>% select(uid, group, duration_aa, duration_as, duration_dd, duration_in, duration_ii)
names(post_times) = c('uid', 'group', 'post_duration_aa', 'post_duration_as', 'post_duration_dd', 'post_duration_in', 'post_duration_ii')
```

merge pre and post

```{r}
pre_post_times <- merge(pre_times, post_times, by= c('uid', 'group'))
```

```{r}
library(ggplot2)
library(tidyr)

# Select relevant columns starting with 'pre_' and 'post_'
variables <- pre_post_times %>%
  select(matches("^pre_|^post_"), group)

# Reshape the data from wide to long format
data_long <- variables %>%
  pivot_longer(cols = -group, names_to = "Variable", values_to = "Value") %>%
  separate(col = "Variable", into = c("Phase", "Variable"), sep = "_", remove = FALSE)

# Create the boxplot grouped by 'group'
#ggplot(df_long, aes(x = Phase, y = duration, fill = group)) +
#  geom_boxplot() +
  #facet_wrap(~ group) +
#  labs(x = "Phase", y = "Duration", title = "Comparison of Pre and Post Durations by Group") +
#    theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

#### show the comparison of time spent on the test between control and experiment for each category

1- AA

```{r}
pre_post_times %>% group_by(group) %>%
  summarize(mean_pre_time_AA = mean(pre_duration_aa),
            sd_pre_time_AA = sd(pre_duration_aa),
            mean_post_time_AA = mean(post_duration_aa),
            sd_post_time_AA = sd(post_duration_aa))
```

2- AS

```{r}
pre_post_times %>% group_by(group) %>%
  summarize(mean_pre_time_AS = mean(pre_duration_as),
            sd_pre_time_AS = sd(pre_duration_as),
            mean_post_time_AS = mean(post_duration_as),
            sd_post_time_AS = sd(post_duration_as))
```

3- DD

```{r}
pre_post_times %>% group_by(group) %>%
  summarize(mean_pre_time_DD = mean(pre_duration_dd),
            sd_pre_time_DD = sd(pre_duration_dd),
            mean_post_time_DD = mean(post_duration_dd),
            sd_post_time_DD = sd(post_duration_dd))
```

4- IN

```{r}
pre_post_times %>% group_by(group) %>%
  summarize(mean_pre_time_IN = mean(pre_duration_in),
            sd_pre_time_IN = sd(pre_duration_in),
            mean_post_time_IN = mean(post_duration_in),
            sd_post_time_IN = sd(post_duration_in))
```

5- II

```{r}
pre_post_times %>% group_by(group) %>%
  summarize(mean_pre_time_II = mean(pre_duration_ii),
            sd_pre_time_II = sd(pre_duration_ii),
            mean_post_time_II = mean(post_duration_ii),
            sd_post_time_II = sd(post_duration_ii))
```

```{r}

pre_post_ng_score %>% filter(!(is.infinite(ng) | is.nan(ng))) %>%
  group_by(group) %>%
  summarize(mean_ng = mean(ng),
            sd_ng = sd(ng))
```

```{r}
library(ggplot2)

ggplot(pre_post_ng_score, aes(x = group, y = ng)) + 
  geom_boxplot()

```

## calculation of Normazlied gain for each category

### 1. Analyzing Arguements

```{r}
aa_pre_post_ng <- pre_post_totals %>% select(uid, group, pre_total_aa, post_total_aa)

aa_pre_post_ng <- aa_pre_post_ng %>% mutate(diff = post_total_aa-pre_total_aa)
aa_pre_post_ng <- aa_pre_post_ng %>% mutate(aa_ng = diff/(8-pre_total_aa))

# aa_pre_post_ng[ is.infinite(aa_pre_post_ng$aa_ng),]$aa_ng = NA

aa_pre_post_ng %>% filter(!(is.infinite(aa_ng) | is.nan(aa_ng))) %>%
  group_by(group) %>%
  summarize(mean_ng = mean(aa_ng),
            sd_ng = sd(aa_ng))
```

### 2. Assumptions

```{r}
as_pre_post_ng <- pre_post_totals %>% select(uid, group, pre_total_as, post_total_as)

as_pre_post_ng <- as_pre_post_ng %>% mutate(diff = post_total_as-pre_total_as)
as_pre_post_ng <- as_pre_post_ng %>% mutate(as_ng = diff/(8-pre_total_as))

as_pre_post_ng %>% filter(!(is.infinite(as_ng) | is.nan(as_ng))) %>%
  group_by(group) %>%
  summarize(mean_ng = mean(as_ng),
            sd_ng = sd(as_ng))
```

```{r}
as_pre_post_ng
```

### 3. Deductions

```{r}
dd_pre_post_ng <- pre_post_totals %>% select(uid, group, pre_total_dd, post_total_dd)

dd_pre_post_ng <- dd_pre_post_ng %>% mutate(diff = post_total_dd-pre_total_dd)
dd_pre_post_ng <- dd_pre_post_ng %>% mutate(dd_ng = diff/(8-pre_total_dd))

dd_pre_post_ng %>% filter(!(is.infinite(dd_ng) | is.nan(dd_ng))) %>%
  group_by(group) %>%
  summarize(mean_ng = mean(dd_ng),
            sd_ng = sd(dd_ng))
```

### 4. inference

```{r}
in_pre_post_ng <- pre_post_totals %>% select(uid, group, pre_total_in, post_total_in)

in_pre_post_ng <- in_pre_post_ng %>% mutate(diff = post_total_in-pre_total_in)
in_pre_post_ng <- in_pre_post_ng %>% mutate(in_ng = diff/(8-pre_total_in))

in_pre_post_ng %>% filter(!(is.infinite(in_ng) | is.nan(in_ng))) %>%
  group_by(group) %>%
  summarize(mean_ng = mean(in_ng),
            sd_ng = sd(in_ng))
```

### 5. interpreting information

```{r}
ii_pre_post_ng <- pre_post_totals %>% select(uid, group, pre_total_ii, post_total_ii)

ii_pre_post_ng <- ii_pre_post_ng %>% mutate(diff = post_total_ii-pre_total_ii)
ii_pre_post_ng <- ii_pre_post_ng %>% mutate(ii_ng = diff/(8-pre_total_ii))

ii_pre_post_ng %>% filter(!(is.infinite(ii_ng) | is.nan(ii_ng))) %>%
  group_by(group) %>%
  summarize(mean_ng = mean(ii_ng),
            sd_ng = sd(ii_ng))
```

#### Analyzing Arguments count plot

```{r}
####  Normalized Change
#count(as_pre_post_ng[as_pre_post_ng$pre_total_as==8 & as_pre_post_ng$post_total_as==8,])
aa_pre_post_ng %>% filter((pre_total_aa==post_total_aa)&(pre_total_aa==8 | pre_total_aa==0))
ggplot(aa_pre_post_ng, aes(x = pre_total_aa, y = post_total_aa)) +
  geom_count(aes(color = ..n.., size = ..n..))
```

### Calculation of standard deviation

```{r}
# Calculate mean and standard deviation of 'score' column grouped by 'group'

pre_post_ng_score %>%
  summarise(mean_total = mean(diff),
            std_dev_total = sd(diff),
            .groups = 'drop') 

pre_post_ng_score %>% filter(abs(diff) < 4) %>%
  summarise(mean_total = mean(diff),
            std_dev_total = sd(diff),
            .groups = 'drop')
#
result <- pre_post_ng_score %>% filter(abs(diff) < 4) %>%
  group_by(group) %>%
  summarise(mean_total = mean(ng),
            std_dev_total = sd(ng),
            .groups = 'drop')

# Print the filtered result
print(result)


```

#### t-test for pre-post-ng between experiment and control group for each category

```{r}
aa_pre_post_ng %>% filter(!is.infinite(aa_ng)) %>% group_by(group) %>% summarize(mean(aa_ng))

t.test(aa_ng ~ group, data = aa_pre_post_ng[!is.infinite(aa_pre_post_ng$aa_ng),], paired=F, var.equal=T)
t.test(as_ng ~ group, data = as_pre_post_ng[!is.infinite(as_pre_post_ng$as_ng),], paired=F, var.equal=T)
t.test(dd_ng ~ group, data = dd_pre_post_ng[!is.infinite(dd_pre_post_ng$dd_ng),], paired=F, var.equal=T)
t.test(in_ng ~ group, data = in_pre_post_ng[!is.infinite(in_pre_post_ng$in_ng),], paired=F, var.equal=T)
t.test(ii_ng ~ group, data = ii_pre_post_ng[!is.infinite(ii_pre_post_ng$ii_ng),], paired=F, var.equal=T)
```

```{r}
library(effsize)
in_df = in_pre_post_ng[!is.infinite(in_pre_post_ng$in_ng),]
# Calculate Cohen's d effect size using a formula
cohen.d(in_df[in_df$group=='exp',]$in_ng, in_df[in_df$group=='cont',]$in_ng, paired = F)
```

```{r}

ii_df = ii_pre_post_ng[!is.infinite(ii_pre_post_ng$ii_ng),]
# Calculate Cohen's d effect size using a formula
cohen.d(ii_df[ii_df$group=='exp',]$ii_ng, ii_df[ii_df$group=='cont',]$ii_ng)
```

#### calculate the total duration for pre and post

```{r}
pre_post_times <- pre_post_times %>% mutate(pre_total_duration = rowSums(select(., starts_with("pre_duration_"))),
                          post_total_duration = rowSums(select(., starts_with("post_duration_"))))
```

### difference in duration on pre-test between the groups

```{r}
t.test(pre_duration_aa ~ group, data = pre_post_times, paired=F, var.equal=T)
t.test(pre_duration_as ~ group, data = pre_post_times, paired=F, var.equal=T)
t.test(pre_duration_dd ~ group, data = pre_post_times, paired=F, var.equal=T)
t.test(pre_duration_in ~ group, data = pre_post_times, paired=F, var.equal=T)
t.test(pre_duration_ii ~ group, data = pre_post_times, paired=F, var.equal=T)
```

#### difference between the groups in post duration

```{r}
t.test(post_duration_aa ~ group, data = pre_post_times, paired=F, var.equal=T)
t.test(post_duration_as ~ group, data = pre_post_times, paired=F, var.equal=T)
t.test(post_duration_dd ~ group, data = pre_post_times, paired=F, var.equal=T)
t.test(post_duration_in ~ group, data = pre_post_times, paired=F, var.equal=T)
t.test(post_duration_ii ~ group, data = pre_post_times, paired=F, var.equal=T)
```

#### difference within the group

-   Experiment group

```{r}
t.test(pre_post_times[pre_post_times$group=='exp',]$post_duration_aa, pre_post_times[pre_post_times$group=='exp',]$pre_duration_aa, paired=T)
t.test(pre_post_times[pre_post_times$group=='exp',]$post_duration_as, pre_post_times[pre_post_times$group=='exp',]$pre_duration_as, paired=T)
t.test(pre_post_times[pre_post_times$group=='exp',]$post_duration_dd, pre_post_times[pre_post_times$group=='exp',]$pre_duration_dd, paired=T)
t.test(pre_post_times[pre_post_times$group=='exp',]$post_duration_in, pre_post_times[pre_post_times$group=='exp',]$pre_duration_in, paired=T)
t.test(pre_post_times[pre_post_times$group=='exp',]$post_duration_ii, pre_post_times[pre_post_times$group=='exp',]$pre_duration_ii, paired=T)
```

-   Control Group

```{r}
t.test(pre_post_times[pre_post_times$group=='cont',]$post_duration_aa, pre_post_times[pre_post_times$group=='cont',]$pre_duration_aa, paired=T)
t.test(pre_post_times[pre_post_times$group=='cont',]$post_duration_as, pre_post_times[pre_post_times$group=='cont',]$pre_duration_as, paired=T)
t.test(pre_post_times[pre_post_times$group=='cont',]$post_duration_dd, pre_post_times[pre_post_times$group=='cont',]$pre_duration_dd, paired=T)
t.test(pre_post_times[pre_post_times$group=='cont',]$post_duration_in, pre_post_times[pre_post_times$group=='cont',]$pre_duration_in, paired=T)
t.test(pre_post_times[pre_post_times$group=='cont',]$post_duration_ii, pre_post_times[pre_post_times$group=='cont',]$pre_duration_ii, paired=T)
```

```{r}
good_students_exp <- aa_pre_post_ng %>% filter(group=='exp' & aa_ng >= -0.46)
good_students_cont <- aa_pre_post_ng %>% filter(group=='cont' & aa_ng >= 0)

summary(good_students_exp$aa_ng)
summary(good_students_cont$aa_ng)

t.test(good_students_exp$aa_ng, good_students_cont$aa_ng, paired = F)
```

```{r}
good_students_exp <- as_pre_post_ng %>% filter(group=='exp' & as_ng >= 0.03)
good_students_cont <- as_pre_post_ng %>% filter(group=='cont' & as_ng >= 0.06)

summary(good_students_exp$as_ng)
summary(good_students_cont$as_ng)

t.test(good_students_exp$as_ng, good_students_cont$as_ng, paired = F)
```

```{r}
good_students_exp <- dd_pre_post_ng %>% filter(group=='exp' & dd_ng >= -0.6)
good_students_cont <- dd_pre_post_ng %>% filter(group=='cont' & dd_ng >= -0.38)

summary(good_students_exp$dd_ng)
summary(good_students_cont$dd_ng)

t.test(good_students_exp$dd_ng, good_students_cont$dd_ng, paired = F)
```

```{r}
good_students_exp <- in_pre_post_ng %>% filter(group=='exp' & in_ng >= -0.07)
good_students_cont <- in_pre_post_ng %>% filter(group=='cont' & in_ng >= -24)

summary(good_students_exp$in_ng)
summary(good_students_cont$in_ng)

t.test(good_students_exp$in_ng, good_students_cont$in_ng, paired = F)
```

```{r}
good_students_exp <- ii_pre_post_ng %>% filter(group=='exp' & ii_ng >= -0.75)
good_students_cont <- ii_pre_post_ng %>% filter(group=='cont' & ii_ng >= -0.09)

summary(good_students_exp$ii_ng)
summary(good_students_cont$ii_ng)

t.test(good_students_exp$ii_ng, good_students_cont$ii_ng, paired = F)
```
